---
title: "TTP Predictive Marker Analysis: Exploratory Analysis"
author: "Chong Kim"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    toc_float: true
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
    #css: my.css   # you can add your custom css, should be in same folder
csl: science.csl
bibliography: bibliography.bib

fontfamily: mathpazo
fontfamilyoptions: sc, osf
fontsize: 11pt
---


<style type="text/css">

TOC {
  position: relative;
  left: 0;
  top: 0;
  width: 200px;
  height: 100%;
  overflow:auto;
}

body {
  max-width: 1200px;
  margin: auto;
  margin-left:210px;
  line-height: 20px;
  font-size: 12px;
}

td {  /* Table  */
  font-size: 13px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

*Please read the [README.md](/https://github.com/ck2136/ttp_pm/blob/master/readme.md) file for more information.*

# Objective

This project is based on a database that was constructed with 73 patients that have Acquired [Thrombotic Thrombocytopenic Purpura (TTP)](https://en.wikipedia.org/wiki/Thrombotic_thrombocytopenic_purpura) at the Unviersity of Alabama Birmingham Hospital. The goal of this project is to determine the predictive markers for outcomes related to TTP patients  antigen levels.

The data have been cleaned but there are a couple of things to note that have been done in the data management process.
1. Repeated measures were not taken into account (i.e. The predictors and outcomes were based ona a cross-sectional measurement).
2. There are only 73 patients and around 58 variables.
3. The outcome of interest is
    A. Mortality
    B. Adverse event (exacerbation, relapse, death) vs. Remission
4. There are many markers that we will see below

# Univariate Analysis

## Table 1: Population Table
```{r ttp_uni_tbl_gen, include=FALSE}
rm(list=ls())
ttpfinal <- readRDS("../data/ttpselected.RDS")


##### Set all the variable
listVars <- names(ttpfinal)

##### Define categorical variables
catVars <- colnames(Filter(is.character, ttpfinal))

##### Overall population table
library(tableone)
library(pander)
table1 <- CreateTableOne(vars = listVars ,data = ttpfinal, factorVars = catVars)
tab1mat <- print(table1, exact = catVars, showAllLevels = TRUE, floating=FALSE,latex.environments=NULL, booktabs=TRUE)
##### tbl2
table2 <- CreateTableOne(vars = listVars, strata = "mort" ,data = ttpfinal, factorVars = catVars)
tab2mat <- print(table2, exact = catVars, showAllLevels = TRUE, floating=FALSE,latex.environments=NULL, booktabs=TRUE)
##### tbl3
table3 <- CreateTableOne(vars = listVars, strata = "outcome2" ,data = ttpfinal, factorVars = catVars)
tab3mat <- print(table3, exact = catVars, showAllLevels = TRUE, floating=FALSE,latex.environments=NULL, booktabs=TRUE)

```

```{r ttp_uni_tbl1}
pander(tab1mat,  split.table = Inf)
```
- The above table is an overall TTP patient summary table.

## Table 2: Stratified by Mortality
```{r ttp_uni_tbl2}
##### Stratified by mortality
pander(tab2mat,  split.table = Inf)
```

- Table 2 is stratified by mortality (1: death; 0: alive). 

## Table 3: Stratified by Adverse Outcomes
```{r ttp_uni_tbl3}
##### Stratified by adverse event
pander(tab3mat,  split.table = Inf)
```

- Table 2 is stratified by adverse event (1/2/3: exacerbation, relapse, or death; 0: remission). 

# Visual Inspection

## Figure 1: 