---
title: "TTP Predictive Marker Analysis: Statistical Analysis Part C"
author: "Chong Kim"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    toc_float: true
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
    #css: my.css   # you can add your custom css, should be in same folder
csl: science.csl
bibliography: bibliography.bib

fontfamily: mathpazo
fontfamilyoptions: sc, osf
fontsize: 11pt
---


<style type="text/css">

TOC {
  position: relative;
  left: 0;
  top: 0;
  width: 200px;
  height: 100%;
  overflow:auto;
}

body {
  max-width: 1200px;
  margin: auto;
  margin-left:210px;
  line-height: 20px;
  font-size: 12px;
}

td {  /* Table  */
  font-size: 13px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

*Please read the [README.md](/https://github.com/ck2136/ttp_pm/blob/master/readme.md) file for more information.*

# Objective

This project is based on a database that was constructed with 73 patients that have Acquired [Thrombotic Thrombocytopenic Purpura (TTP)](https://en.wikipedia.org/wiki/Thrombotic_thrombocytopenic_purpura) at the Unviersity of Alabama Birmingham Hospital. The goal of this project is to determine the predictive markers for outcomes related to TTP patients  antigen levels.

The previous [exploratory analysis](ttp_pm_report1.html) indicated some significant univariate association between predictors and mortality (Other outcomes such as exacerbation and relapse didn't show any strong association with predictors in the data). [Statistical Analysis A](ttp_pm_report2.html) and  [Statistical Analysis B](ttp_pm_report2-b.html) approach the analysis using machine learning algorithms.

In this report, we will go through analyzing the time to platelet stabilization using a survival analysis framework  (e.g. Cox proportional hazard regression) to determine which factors are highly associated with risk of platelet non-stabilization(or stabilization) in TTP patients.

# Exploratory Visualization 

## Kaplan Meier Curve

- This is just an illustration to indicate:
    - Probability of non-stabilization (given our outcome here is time to stabilization)
    - Median time to stabilization
    - Potential Follow-up time
    - Visual inspection of hazard function over time

### Table 1: logistic regression with all variabiles (based on univariate p-value)
```{r ttp_km}
library(dplyr)
library(magrittr)
rm(list=ls())
ttpfinal <- readRDS("~/Documents/Projects/Huy/ttp_pm/data/ttpselected.RDS")
ttpfinalcat <- readRDS("../data/ttpselectedcat.RDS")

merged <- ttpfinal %>%
  mutate(id = seq(1:nrow(.))) %>%
  left_join(
    ttpfinalcat %>%
      select(which(sapply(ttpfinalcat, class) == 'character')) %>%
      set_colnames(., unlist(sapply(names(ttpfinalcat %>% 
                                            select(which(sapply(ttpfinalcat, class) == 'character'))), function(x) {
                                              paste0(x,"_cat") 
                                            }))) %>%
      mutate(id = seq(1:nrow(.))) %>%
      select(-c(1:2)), by = "id"
  ) %>%
  mutate(race = ifelse(race == "Afro-Hispanic", "AA", race),
         plt_stb_7 = as.character(plt_stb_7))
  
temp <- merged
temp$mort <- as.numeric(temp$mort)
# kaplan-meier calculation (non parametric survival probabilities)
library(survival)
result.km <- survfit(Surv(temp$los, temp$mort) ~ 1, conf.type = "log-log")
# To calculated the potential follow-up time
temp$fu <- 1 - temp$mort
result.rkm <- survfit(Surv(temp$los, temp$fu) ~ 1, conf.type = "log-log")

library(survminer)

# Customized survival curves
ggsurvplot(result.km, data = temp,
           surv.median.line = "hv", # Add medians survival
           
           # Change legends: title & labels
           #legend.title = "Sex",
           #legend.labs = c("Male", "Female"),
           # Add p-value and tervals
           #pval = TRUE,
           
           conf.int = TRUE,
           # Add risk table
           risk.table = TRUE,
           tables.height = 0.2,
           tables.theme = theme_cleantable(),
           
           # Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
           # or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
           #palette = c("#E7B800", "#2E9FDF"),
           ggtheme = theme_bw() # Change ggplot2 theme
)
```

The median time to survival is `r summary(result.km)$table[[7]]` days. 

```{r exp_haz_func, include = FALSE}
library(muhaz)
# we want to check the hazard at 5 and 10 months interval
result.pe10 <- pehaz(temp$los, temp$mort, width = 10, 
                     max.time = 60)
result.pe5 <- pehaz(temp$los, temp$mort, width = 5, 
                     max.time = 60)

result.smooth <- muhaz(temp$los, temp$mort, bw.smooth = 10, 
                       b.cor = "left", max.time = 60)

```


```{r exp_haz_func_plot}
plot(result.pe10, ylim = c(0, 0.3), col = "blacK")
title("Hazard Function Estimation")
lines(result.pe5, col = "green")
lines(result.smooth, col = "red")
```

we can clearly see a decreasing hazard function... in this case an exponential would not be appropriate. 

# Univariate Analysis

## Cox-PH regression Significant Variables

Below are the univariately statistically significant variables based on modeling a Cox-PH regression for continuous variables.

```{r univ_coxph}

temp_cont <- temp %>%
  select(which(sapply(temp, class) != 'character'), -outcome2, los, mort)

# fit univariate survival models to all variables and check wald-test for fit
tmp <- sapply(setdiff(colnames(temp_cont), c("time","status","los","mort","fum","eculizumab", "cyclophosphamide", "transplant", "outcome1","outcome2","outcome2_cat")), function(x) {
  try(
    c(summary(coxph(Surv(los, mort) ~ get(x), data = temp_cont))$waldtest[[3]] < 0.05,
      round(summary(coxph(Surv(los, mort) ~ get(x), data = temp_cont))$waldtest[[3]], 3),
      round(exp(confint(coxph(Surv(los,mort) ~ get(x), data = temp_cont))[1]), 3),
      round(exp(confint(coxph(Surv(los,mort) ~ get(x), data = temp_cont))[2]), 3),
      round(summary(coxph(Surv(los,mort) ~ get(x), data = temp_cont))$coefficients[2], 3)
    ), TRUE)
})


# to select p-values based on variables that had p <= 0.05
library(dplyr)
library(magrittr)
tmp1 <- tmp %>%
  t(.) %>%
  set_colnames(c("Sign","p","LCI","UCI","Coef"))  %>%
  as.data.frame() %>%
  cbind(rownames(.)) %>%
  set_colnames(c("Sign","p","LCI","UCI","Coef","Varn"))  %>%
  select(p, LCI, UCI, Coef, Varn) %>%
  mutate(p = as.numeric(p)) %>%
  filter(., p <= 0.05)

library(pander)
pander(tmp1, split.table = Inf)
```

Below are the univariately statistically significant variables based on modeling a Cox-PH regression for categorical variables.


```{r univ_coxph_cat}

temp_cat <- temp %>% select(which(sapply(temp, class) == 'character'), -outcome1,-outcome2,-outcome2_cat, los, mort)

tmp_cat <- sapply(setdiff(colnames(temp_cat), c("time","status","los","mort","fum","eculizumab", "cyclophosphamide", "transplant", "outcome1", "outcome2")), function(x) {
  try(
    c(summary(coxph(Surv(los, mort) ~ get(x), data = temp_cat))$waldtest[[3]] < 0.05,
      round(summary(coxph(Surv(los, mort) ~ get(x), data = temp_cat))$waldtest[[3]], 3)
      ), TRUE)
})

tmp2 <- tmp_cat %>%
  t(.) %>%
  set_colnames(c("Sign","p"))  %>%
  as.data.frame() %>%
  cbind(rownames(.)) %>%
  set_colnames(c("Sign","p", "Varn"))  %>%
  select(p, Varn) %>%
  mutate(p = as.numeric(p)) %>%
  filter(., p <= 0.05)

pander(tmp2, split.table = Inf)
```

unfortunately if we check the regression estimates (i.e. standard error and etc) we notice that the significance is due to problems with how the pai_cat variable is distributed  across the two groups. (i.e. small sample size).

```{r pai_prob, results="asis"}
library(stargazer)
stargazer(coxph(Surv(los, mort) ~ pai_cat, data = temp_cat), ci = TRUE, type = "html")
```